### 2024-11-14 docs/manuscript.tex
Updated manuscript with reviewer suggestions.

@@

The primary objective of this study is to examine whether the academic
department affiliation of a corresponding author has a discernible
outcome on the accuracy (i.e. correctness of predictions) of the
bioinformatic tools they develop. Specifically, we aim to determine
whether tools created by authors from domain-expert,
development-expert of interdisciplinary fields differ in accuracy. To
address this, we analyzed a benchmarked corpus of bioinformatic
software tools and evaluated their accuracy based on the developers'
academic affiliations.

\begin{figure*}[t!]
  \centering
  \includegraphics[width=0.75\textwidth]{upset-plots.pdf}
  \caption{ The number and intersection of general (\textbf{A}) or
  specific (\textbf{B}) fields that have contributed to bioinformatic
  software tools included in recent benchmark studies.  The number and
  intersection of specific fields.  }
\label{fig:fig1}
\end{figure*}

@@

We explored the relationship between the accuracy of bioinformatic
software tools and the academic fields of their developers. Using a
previously published corpus of benchmarked accuracy rankings
\cite{gardner2024}, we mapped corresponding authors' addresses to
standardized ``fields of study'' \cite{fields2014} and grouped them into
broader categories.

%
Figure~\ref{fig:fig1} shows the number of tools for the general and
specific fields ($N\ge 10$). Most bioinformatic tools were developed
by authors affiliated with Genetics, Bioinformatics, Computer Science,
or similar departments. Among the general fields, Biological Sciences
produced the most software tools, followed by Computer Sciences.

\begin{figure*}[ht!]
\begin{center}
  \includegraphics[width=0.65\textwidth]{forest-z-Plot.pdf}
\end{center}
\caption{(\textbf{A}) A forest plot, illustrating the mean and 95\%
  confidence intervals of the proportion of times software tools
  published by a given field ``win'' in pairwise
  comparisons. Confidence intervals and the mean was determined using
  a bootstrapping procedure. Within each field the entries have been
  sorted by the mean number of wins. The sample size for each field is
  indicated by the column of numbers on the right of the figure.
  (\textbf{B}) A Z-score was computed for each distribution of
  bootstrap samples for each field. The expected proportion of wins
  for randomly selected groups of tools was used as ``$x$''
  (i.e. null=0.5).}
\label{fig:fig2}
\end{figure*}

@@

\textbf{Benchmarking data:} software ranks from previously gathered
benchmarks are publically available \cite{Gardner:2022}, these include data from
68 publications that rank the accuracy of different sets of 498
distinct software tools.

@@

>>>>

%\definecolor{rv}{RGB}{0,0,200} %colour highlighting revisions
\definecolor{rv}{RGB}{0,0,0} %colour masking revisions

@@

\textcolor{rv}{The accuracy of bioinformatic software tools are of
  critical importance to the field \cite{weber2019essential}. Yet
  several factors can influence accuracy assessments. Previous work
  has shown that author conflicts can bias tool accuracy assessments
  upwards \cite{buchka2021optimistic}. Conversely, commonly assumed
  features of accurate software such as author reputation
  (e.g. H-index), tool citation rates, journal impact, runtime, and
  age of tools had no significant association with software accuracy
  \cite{Gardner:2022}.  Instead, features illustrating sustained
  development of tools, such as the number of releases, and activity
  on a popular distributed version control website was significantly
  associated with accuracy \cite{Gardner:2022}.}

The primary objective of this study is to examine whether the academic
department affiliation of a corresponding author has a discernible
outcome on the accuracy (i.e. correctness of predictions) of the
bioinformatic tools they develop. Specifically, we aim to determine
whether tools created by authors from domain-expert,
development-expert of interdisciplinary fields differ in accuracy.
\textcolor{rv}{This question was motivated by the possibility
  that certain grant reviewers and other research assessors may base
  their perception of expertise on departmental affiliation rather
  than more pertinent information such as research experience.} To
address this challenge we have analyzed a published corpus of
benchmarked bioinformatic tools and evaluated relative accuracy ranks
based on the developers' academic affiliations.

\begin{figure*}[t!]
  \centering
  \includegraphics[width=0.75\textwidth]{figure1.pdf}
  \caption{ \textcolor{rv}{ The figure displays the number and
      intersection of bioinformatic tools classified by general fields (Panel \textbf{A})
      and specific fields (Panel \textbf{B}), as listed by the corresponding
      authors who published software included in benchmark
      studies. Panel \textbf{C} presents the classification of general and
      specific fields of study used by the ``National Science Foundation''
      relevant to this study. On the right, it shows the additional
      expertise categories that we introduced for this analysis.
 }  }
\label{fig:fig1}
\end{figure*}

@@

We explored the relationship between the accuracy of bioinformatic
software tools and the academic fields of their developers. Using a
previously published corpus of benchmarked accuracy rankings
\cite{gardner2024}, we mapped corresponding authors' addresses to
standardized ``fields of study'' \cite{fields2014} and grouped them
into broader categories. Fields of study is a hierarchical
classification that includes specific fields (e.g. ``Genetics'' or
``Mathematics'') or general fields (e.g. ``Biological Sciences'' or
``Mathematics and Statistics''). We further classified these into
expertise areas of ``Development'' (e.g. Computer Science), ``Domain''
(e.g. Biological Sciences) and ``Interdisciplinary''
(e.g. Bioinformatics or listing both a Biological and Computational
Sciences affiliation).

%
\textcolor{rv}{Figure~\ref{fig:fig1} shows the number of tools (set/intersection sizes) for the general
and specific fields (for sets sizes $\ge 10$), as well as the intersections between
fields due to corresponding authors with multiple
affiliations. E.g. 11 tools have been published by authors who list
both a Genetics and Computer Science department as their
affiliation.} Most bioinformatic tools were developed by authors
affiliated with Genetics, Bioinformatics, Computer Science, or similar
departments. Among the general fields, Biological Sciences produced
the most software tools, followed by Computer Sciences.

\begin{figure*}[ht!]
\begin{center}
  \includegraphics[width=0.65\textwidth]{figure2.pdf}
\end{center}
\caption{(\textbf{A}) A forest plot, illustrating the mean and 95\%
  confidence intervals of the proportion of times software tools
  published by a given field ``win'' in pairwise
  comparisons. Confidence intervals and the mean was determined using
  a bootstrapping procedure. Within each field the entries have been
  sorted by the mean number of wins. The sample size for each field is
  indicated by the column of numbers on the right of the figure.
  (\textbf{B}) A Z-score was computed for each distribution of
  bootstrap samples for each field. The expected proportion of wins
  for randomly selected groups of tools was used as ``$x$''
  (i.e. null=0.5).}
\label{fig:fig2}
\end{figure*}

@@

\textbf{Benchmarking data:} software ranks from previously gathered
benchmarks are publically available \cite{Gardner:2022}, these include data from
68 publications, \textcolor{rv}{128 benchmark rankings of} different sets of 498
distinct software tools (See Supplementary Table 1).

\textcolor{rv}{In brief, the criteria for inclusion of a
  benchmark study are based on the definition of ``neutral comparison
  studies'' \cite{boulesteix2013plea}: 1. The main focus of the
  article is the comparison and not the introduction of a new
  tool. 2. The authors should be neutral (i.e. not tool
  developers). 3. The test data and evaluation criteria should be
  sensible.}

@@

\textcolor{rv}{The author profusely thanks the recent anonymous grant
  reviewers who kindly noted the author’s
  bioinformatics/machine-learning skills as 'weak' or 'lacking
  expertise'. We assume they were swayed more by our affiliation than
  by minor details like... a publication record. This charming
  oversight, of course, inspired the paper you're reading now.}
